---
title: "cRacle_BETA"
author: "Rob Harbert"
date: "8/17/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(cRacle)
```

## Announcing a new R package dedicated to implementing CRACLE for climate estimation from biological community data.

## How to:

First load some climate data:

```{r}
data(climondbioclim) #prepackaged with cRacle
#advise:
#clim = raster::getData('worldclim', var='bio', res=2.5)
```
For a list of species:

```{r}

sp_list = c(
  "Quercus rubra",
  "Pinus strobus",
  "Vaccinium angustifolium",
  "Betula papyrifera",
  "Rhododendron maximum",
  "Toxicodendron radicans"
)

```

Get data from GBIF.
```{r}

extr = getextr(
  sp_list,
  climondbioclim,
  repo=c('gbif', 'inat'),
  maxrec = 200,
  #Should set higher in practice! 
  schema = 'flat',
  factor = 2,
  #adjust to spatially thin. Use a higher number (4 to 8) for spatial grids of 2.5 arcminutes or less.
  nmin = 10,
  parallel = FALSE
)
```

Then build probability distributions for each taxon/variable pair.

```{r}
densall = dens_obj(extr,
                   climondbioclim,
                   manip = 'condi',
                   kern = 'gaussian',
                   bg.n = 40) #consider using parallel=TRUE if bg.n>500
```

Then calculate the intersect of all taxa for each variable. This is the Coexistence Likelihood Estimation.

```{r}
and = and_fun(densall)

```

And find the optima for these joint likelihood functions.
```{r}
optim = get_optim(and)

print(optim$origk) #As implemented in Harbert and Nixon, 2015
print(optim$conintkde) #with 95% confidence intervals on the likelihood distribution.
```

And visualize (for Mean Annual Temperature):
```{r multiplot, echo =TRUE}
n = 1
multiplot(densall, names(climondbioclim)[[n]])
addplot(and, names(climondbioclim)[[n]])

abline(v = median(optim$origk[, n]), col = 'green') #Midpoint will be very close to the optimum. Original method of Harbert and Nixon, 2015 returned a top 1% range to introduce some slop

```
